\chapter{Generating Uniform Random Numbers}

\setcounter{problem}{1}

\begin{quote}
{\bf Q}: {\em How to generate pure random string}? \\
{\bf A}: {\em Put a fresh student in front of vi editor and ask him to quit}.\\
~~~~--- Emiliano Lourbet (@taitooz)
\end{quote}

\section{Discussion}

\begin{fullwidth}

To perform computer-based simulation we need to be able to generate random numbers. Generating random numbers following a uniform distribution are the easiest to generate and are what comes out of the standard programming language ``give me a random number'' function.  Here's a sample Python session:

\begin{pyconsole}
import random
print random.random()
print random.random()
print random.random()
\end{pyconsole}

We could generate real random numbers by accessing, for example, noise on the ethernet network device but that noise might not be uniformly distributed. We typically generate pseudorandom numbers that aren't really random but look like they are. From Ross' {\em Simulation},  we see a very easy recursive mechanism that generators values in $[0,m-1]$:\\

$x_n = a x_{n-1}$ modulo $m$\\

That's recursive (or iterative and not {\em closed form}) because $x_n$ is a function of a prior value: \\

$x_1 = ax_0$ modulo $m$\\
$x_2 = ax_1$ modulo $m$\\
$x_3 = ax_2$ modulo $m$\\
$x_4 = ax_3$ modulo $m$\\
$...$\\

\noindent To get random numbers between 0 and 1, we use $x_n / m$.

We must pick a value for $a$ and $m$ that make $x_n$ seem random. Ross suggests choosing a large prime number for $m$ that fits in our integer word size, e.g., $m = 2^{31} - 1$, and $a = 7^5 = 16807$.

Initially we set a value for $x_0$, called the {\em random seed} (it is not the first random number). Every seed leads to a different sequence of pseudorandom numbers. In Python, you can set the seed of the standard library by using {\tt random.seed([x])}.

Your goal is to take that simple recursive formula and use it to generate the first 10 random numbers using a {\tt for} loop in Python as part of the ``main'' code. Please use file {\tt varunif.py} and make function {\tt runif()} that returns a new random value per call. Use $m = 2^{31} - 1$, $a = 7^5 = 16807$, and an initial seed of $x_0 = 666$.  Your output should look something like:

\begin{verbatim}
0.00521236192678
0.604166903349
0.233144581892
0.460987861017
0.822980116505
0.826818094508
0.331714398848
0.1239014343
0.411406287184
0.505468696591
\end{verbatim}

Because we are all using the same seed, the sequence of numbers should be the same.

Next, make function {\tt runif\_(a,b)} that returns random values between {\tt a} and {\tt b}. Your function definitions should look like:

\begin{pyverbatim}
# U(0,1)
def runif():
    ...

# U(a,b)
def runif_(a,b):
    ...
\end{pyverbatim}

\section{Deliverables}

You must submit your {\tt varunif.py} containing a main loop and your functions runif and runif\_. Also submit your output via Canvas as a text file.

\end{fullwidth}
